A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-05 docs/methods/erosion_methods.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  There are several models for
erosion~\cite{par-izu2000,lag2000}, but we have adopted the model that
erosion occurs in the normal direction at a rate that is proportional to
the magnitude of the shear stress~\cite{moo-ris-chi-zha-she2013}.  The
constant of proportionality only sets the time scale, so we have set
$\Vn = |\tau|$ as described in equation~\eqref{eqn:erosionModel}.  For
the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Formulation}
\label{s:formulation}
We start by defining the main variables used to model erosion.  We
consider flows that are confined by a solid wall $\Gamma$ that encloses
$M$ eroding bodies as diagrammed in Fig.~\ref{fig:schematic}.  The
bodies are denoted as $\gamma_\ell$, $\ell=1,\ldots,M$, and we write
$\gamma = \gamma_1 \cup \cdots \cup \gamma_M$.  On each body, $\nn$ and
$\ss$ denote the unit normal and unit tangent vectors respectively.
Throughout the paper, we adopt the convention that $\ss$ points in the
counterclockwise direction and $\nn$ points out of the fluid domain
(i.e.~into the bodies). Neglecting inertial forces, the dynamics of the
fluid is fully determined by the position of the bodies $\xx_\ell(s,t)
\in \gamma_\ell$, where $s$ is the arclength and $t$ is time.  Given
$\xx_\ell(s,t)$, $\ell=1,\ldots,M$, derived variables include the fluid
velocity $\uu$, the pressure $p$, and the shear stress $\tau$.   On the
bounding wall, we prescribe a velocity $\UU(\xx,t)$.  Then, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
\begin{split}
  \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap &&\mbox{conservation
of momentum}\\
\grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap
&&\mbox{\em conservation of mass} \\
\uu = 0, &\hspace{20pt} \xx \in \gamma, \gap &&\mbox{\em no slip on the
bodies} \\
\uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap &&\mbox{\em outer wall
velocity} \\
\Vn = \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
&&\mbox{\em erosion model},
\end{split}
\end{equation}
where $\Vn$ is the normal velocity of the interface and $\mu$ is the
fluid viscosity.  There are several models for
erosion~\cite{par-izu2000,lag2000}, but we have adopted the model that
erosion occurs in the normal direction at a rate that is proportional to
the magnitude of the shear stress~\cite{moo-ris-chi-zha-she2013}.  
\todo[inline]{This makes it sound the erosion law sound too will-nilly; like anyone can come up with their own erosion law tomorrow and it would be equally valid. This is not the case. We use the law in which erosion is proportional to shear stress because {\em that is what the laboratory experiments show is true} (Ristoph 2012).}

The constant of proportionality only sets the time scale, so we have set
$\Vn = |\tau|$ as described in equation~\eqref{eqn:erosionModel}.  For
the sake of numerical stability, we modify this law in
Section~\ref{sec:thetaL}.
